\sys: Light-touch Data-Parallel Shell Processing

### 2020-10-09 paper/draft.tex
Corrections

\TODO{Add pointers}

\heading{Commands}
To understand standard commands available in any shell, \sys groups POSIX and GNU commands into a small but well-defined set of \emph{parallelizability classes}.
Rather than describing a command's full observable behavior, these classes focus on information that is important for data parallelism.
To allow other commands to use its transformations, \sys defines a light \emph{annotation language} for describing a command's parallelizability class.
Annotations are expressed once per command rather than once per script and aim towards command developers rather than its users, so that they can quickly and easily capture the characteristics of the commands they develop.
% We haven't introduced the notion of flags
% A command developer can only annotate the default behavior;
%   if a new flag is added a few years later, and if that flag alters the command's class, only then a new annotation is needed.

@@

\heading{Scripts}
To maintain sequential semantics, \sys first analyzes a script to identify \emph{parallelizable regions} containing commands that are candidates for parallelization.
This analysis is guided by the script structure: % to identify parallelization candidates:
  some constructs expose task parallelism; others enforce synchronization.
% \sys first needs to understand the POSIX constructs used to compose larger programs, indicating different levels of candidacy for pararllelization.
%% To assign a candidacy for parallelization, \sys breaks POSIX constructs into a few broad classes:
%%  (i) the parallel-composition operator (\ttt{\&}) exposes task parallelism,
%%  (ii) special commands such as \ttt{xargs} and \ttt{tee} expose data parallelism,
%%  (iii) the pipe operator (\ttt{|}) is a candidate for both task and data parallelism, and
%%  (iv)  other composition operators such as \ttt{\&\&}, \ttt{||}, and \ttt{;} are barrier candidates.
% Since the shell representation only exposes coarse grained parallelization opportunities, 
\sys then converts each parallelizable region to a \emph{dataflow graph} (DFG), 
%% Provided information commands and POSIX constructs, \sys needs to analyze the given script and identify program fragments that are candidates for parallelization.
%% To achieve this, it converts sections of the script that are amenable to parallelization into a data-flow graph (DFG) representation.
% Second, \sys defines a dataflow graph model that is suitable for the semantics of the POSIX shell and proposes a set of parallelization optimizations that preserve the behavior of the dataflow graph.
  a flexible representation that enables a series of local transformations to expose data parallelism, converting the graph into its parallel equivalent.
%% , in which DFGs indicate the parallelizable regions.
Further transformations compile the DFG back to a shell script that uses POSIX constructs to guide parallelism explicitly while aiming at preserving the semantics of the sequential program.

\heading{Implementation}
\sys addresses several practical challenges through a set of constructs it provides---\eg a set of
%% linkable
modular components for augmenting command composition.
%% for stream splitting and merging.
It also provides a small library of efficient data aggregators for a large set of parallelizable commands.
%% For the vast majority of commands in one of the parallelizable classes, it also provides  a small library of pure fold function implementations.
%% \kk{The terms pure, classes, and fold don't mean a lot to the reader at the moment, so I would prefer using more general terminology.}
All these commands live in the \ttt{PATH} and are addressable by name, which means they can be used like (and by) any other commands.

@@

For example, consider \ttt{comm}, which performs a join-like operation on its two inputs to identify common elements.
When \ttt{comm} is invoked without any flags, it produces a three-column output:
  the first column contains lines that are unique to the first input, the second column contains lines that are unique to the second input, and the last column contains lines that exist in both inputs.
A user can invoke \ttt{comm} with any combination of flags \ttt{-1}, \ttt{-2}, or \ttt{-3} to suppress the corresponding output column(s).
% The first clause from its annotation record is shown below:
Its annotation record is shown on the left:
%
% \begin{minted}[fontsize=\footnotesize]{json-object}
%   "predicate":
%     { 
%       "operator": "all",
%       "operands": ["-1", "-3"] 
%     },    
%   "class": "stateless",
%   "inputs": ["args[1]"],
%   "outputs": ["stdout"],
%   "options": ["stdin-hyphen"]
% \end{minted}

@@

\noindent
As the first and third columns are suppressed, \ttt{comm} can be categorized in \sta by regarding its first file argument as a static ``configuration'' input;
% NV: I don't think this is correct, because the small input might be the other argument;  you could even have `comm - -`
  this input accompanies every replicated instance of the command executing in parallel. The second clause is symmetric to the first with respect to the second input, and the third clause, which is the general case classifies \ttt{comm} as \pur as it needs to read both inputs completely before outputing its results. The option \ttt{stdin-hyphen} indicates that \ttt{-} represents standard input if given as a file argument.
% The third clause is the general case, which includes \ttt{comm} invocations without any options, and is captured by the third clause.
% There, \ttt{comm} is classified as \pur as it needs to % wait until it 
%   read both inputs completely before it can output its results.
% Its inputs are its first and second file arguments (\ie excluding any options), and its output is produced in \ttt{stdout}.

@@

\subsection{DFG Definitions}
\label{graph-components}

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{\detokenize{./figs/tiling_throughput_scaleup.pdf}}
    \caption{
      \textbf{\sys's speedup for \ttiny{width=} 2--64.}
      Different configurations per benchmark:
      (i) PaSh: the complete implementation with \ttiny{eager} and \ttiny{split} enabled,
      % (ii) Par+B.Split: \ttt{eager} and input-aware \ttt{split} enabled,
      (ii) PaSh w/o split: \ttiny{eager} enabled (no \ttiny{split}),
      (iii) \todo{Blocking Eager}, only blocking \ttiny{eager} enabled (no \ttiny{split}),
      (iv) \todo{No Eager}: both \ttiny{eager} and \ttiny{split} disabled.
  For some pairs of configurations, \sys produces identical parallel scripts and thus only one is shown.
}
    \vspace{-15pt}
    \label{fig:microbenchmarks}
\end{figure*}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{\detokenize{./figs/unix50_individual_speedups_16.pdf}}
  \caption{
	  \textbf{Unix50 scripts.}
    Speedup (left axis) over sequential execution (right axis) for Unix50 scripts.
    Parallelism is 16$\times$ on 10GB of input data~\cf{unix50}.
    Pipelines are sorted in descending speedup order. % \str
    % \kk{FUTURE: Add speedup by primitives and transformations in bar-chart.}
    % The solid line indicates the average speedup $\times5.49$ and the dotted line indicates 1.
    % \nv{Idea: scripts that only take over 0.01s?}
    % \kk{Now that we removed the small ones, sequential time does not add a lot to this plot. Maybe we can just show speedups?}
  }
  \vspace{-15pt}
  \label{fig:unix50-individual}
\end{figure}

>>>>

\heading{Commands}
To understand standard commands available in any shell, \sys groups POSIX and GNU commands into a small but well-defined set of \emph{parallelizability classes}~\sx{cmd}.
Rather than describing a command's full observable behavior, these classes focus on information that is important for data parallelism.
To allow other commands to use its transformations, \sys defines a light \emph{annotation language} for describing a command's parallelizability class~\sx{ext}.
Annotations are expressed once per command rather than once per script and aim towards command developers rather than its users, so that they can quickly and easily capture the characteristics of the commands they develop.
% We haven't introduced the notion of flags
% A command developer can only annotate the default behavior;
%   if a new flag is added a few years later, and if that flag alters the command's class, only then a new annotation is needed.

@@

\heading{Scripts}
To maintain sequential semantics, \sys first analyzes a script to identify \emph{parallelizable regions} containing commands that are candidates for parallelization~\sx{dataflow-regions}.
This analysis is guided by the script structure: % to identify parallelization candidates:
  some constructs expose task parallelism; others enforce synchronization.
% \sys first needs to understand the POSIX constructs used to compose larger programs, indicating different levels of candidacy for pararllelization.
%% To assign a candidacy for parallelization, \sys breaks POSIX constructs into a few broad classes:
%%  (i) the parallel-composition operator (\ttt{\&}) exposes task parallelism,
%%  (ii) special commands such as \ttt{xargs} and \ttt{tee} expose data parallelism,
%%  (iii) the pipe operator (\ttt{|}) is a candidate for both task and data parallelism, and
%%  (iv)  other composition operators such as \ttt{\&\&}, \ttt{||}, and \ttt{;} are barrier candidates.
% Since the shell representation only exposes coarse grained parallelization opportunities, 
\sys then converts each parallelizable region to a \emph{dataflow graph} (DFG)~\sx{graph-components}, 
%% Provided information commands and POSIX constructs, \sys needs to analyze the given script and identify program fragments that are candidates for parallelization.
%% To achieve this, it converts sections of the script that are amenable to parallelization into a data-flow graph (DFG) representation.
% Second, \sys defines a dataflow graph model that is suitable for the semantics of the POSIX shell and proposes a set of parallelization optimizations that preserve the behavior of the dataflow graph.
  a flexible representation that enables a series of local transformations to expose data parallelism, converting the graph into its parallel equivalent~\sx{ir:transformations}.
%% , in which DFGs indicate the parallelizable regions.
Further transformations compile the DFG back to a shell script that uses POSIX constructs to guide parallelism explicitly while aiming at preserving the semantics of the sequential program~\sx{backend}.

\heading{Implementation}
\sys addresses several practical challenges through a set of constructs it provides---\ie 
%% linkable
modular components for augmenting command composition~\sx{impl}.
%% for stream splitting and merging.
It also provides a small and efficient \emph{aggregator library} targeting a large set of parallelizable commands.
%% For the vast majority of commands in one of the parallelizable classes, it also provides  a small library of pure fold function implementations.
%% \kk{The terms pure, classes, and fold don't mean a lot to the reader at the moment, so I would prefer using more general terminology.}
All these commands live in the \ttt{PATH} and are addressable by name, which means they can be used like (and by) any other commands.

@@

For example, consider \ttt{comm}, which performs a join-like operation on its two inputs to identify common elements.
When \ttt{comm} is invoked without any flags, it produces a three-column output:
  the first column contains lines that are unique to the first input, the second column contains lines that are unique to the second input, and the last column contains lines that exist in both inputs.
A user can invoke \ttt{comm} with any combination of flags \ttt{-1}, \ttt{-2}, or \ttt{-3} to suppress the corresponding output column(s).
% The first clause from its annotation record is shown below:
Below is its annotation record (left) and the implementation of record's first clause (right):
%
% \begin{minted}[fontsize=\footnotesize]{json-object}
%   "predicate":
%     { 
%       "operator": "all",
%       "operands": ["-1", "-3"] 
%     },    
%   "class": "stateless",
%   "inputs": ["args[1]"],
%   "outputs": ["stdout"],
%   "options": ["stdin-hyphen"]
% \end{minted}

@@

\noindent
As the first and third columns are suppressed, \ttt{comm} can be categorized in \sta by regarding its first file argument as a static ``configuration'' input;
% NV: I don't think this is correct, because the small input might be the other argument;  you could even have `comm - -`
  this input accompanies every replicated instance of the command executing in parallel.
The second clause is symmetric to the first with respect to the second input. 
The third clause is the general case, classifiying \ttt{comm} in \pur as it needs to read both inputs completely before outputing its results.
Option \ttt{stdin-hyphen} indicates that \ttt{-} represents standard input if given as a file argument.
% The third clause is the general case, which includes \ttt{comm} invocations without any options, and is captured by the third clause.
% There, \ttt{comm} is classified as \pur as it needs to % wait until it 
%   read both inputs completely before it can output its results.
% Its inputs are its first and second file arguments (\ie excluding any options), and its output is produced in \ttt{stdout}.

@@

\subsection{Dataflow Model Definitions}
\label{graph-components}

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{\detokenize{./figs/tiling_throughput_scaleup.pdf}}
    \vspace{-20pt}
    \caption{
      \textbf{\sys's speedup for \ttiny{width=}2--64.}
      Different configurations per benchmark:
      (i) PaSh: the complete implementation with \ttiny{eager} and \ttiny{split} enabled,
      % (ii) Par+B.Split: \ttt{eager} and input-aware \ttt{split} enabled,
      (ii) PaSh w/o split: \ttiny{eager} enabled (no \ttiny{split}),
      (iii) \todo{Blocking Eager}, only blocking \ttiny{eager} enabled (no \ttiny{split}),
      (iv) \todo{No Eager}: both \ttiny{eager} and \ttiny{split} disabled.
  For some pairs of configurations, \sys produces identical parallel scripts and thus only one is shown.
}
    \vspace{-14pt}
    \label{fig:microbenchmarks}
\end{figure*}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{\detokenize{./figs/unix50_individual_speedups_16.pdf}}
  \vspace{-15pt}
  \caption{
	  \textbf{Unix50 scripts.}
    Speedup (left axis) over sequential execution (right axis) for Unix50 scripts.
    Parallelism is 16$\times$ on 10GB of input data~\cf{unix50}.
    Pipelines are sorted in descending speedup order. % \str
    % \kk{FUTURE: Add speedup by primitives and transformations in bar-chart.}
    % The solid line indicates the average speedup $\times5.49$ and the dotted line indicates 1.
    % \nv{Idea: scripts that only take over 0.01s?}
    % \kk{Now that we removed the small ones, sequential time does not add a lot to this plot. Maybe we can just show speedups?}
  }
  \vspace{-15pt}
  \label{fig:unix50-individual}
\end{figure}
